---
title: 'Chapter 2: Data'
author: "Paige Bailey"
date: "April 8, 2017"
output:
  html_document:
    theme: "sandstone"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-----

## Question 1
In the initial example of Chapter 2, the statistician says "Yes, fields 2 and 3 are basically the same." Can you tell from the three lines of sample data that are shown why she says that?

Field 1 | Field 2 | Field 3 | Field 4 | Field 5
--- | --- | --- | --- | ---
012 | 232 | 33.5 | 0 | 10.7
020 | 121 | 16.9 | 2 | 210.1
027 | 165 | 24.0 | 0 | 427.6

##### **Answer:**

Upon inspection of the table above, we can see that the second column (containing the values 232, 121, and 165) divided by $\approx$ 7 would create the values in the third column (33.5, 16.9, 24.0).

``` {r}
field_2 <- c(232, 121, 165)
field_3 <- c(33.5, 16.9, 24.0)
field_2 / field_3
```

-----

## Question 2
Classify the following attributes as _binary_, _discrete_, or _continuous_. Also classify them as qualitative (**nominal** or **ordinal**) or quantitative (**interval** or **ratio**). Some cases may have more than one interpretation, so briefly indicate your reasoning if you think there may be some ambiguity.

A) Time in terms of AM or PM. 
B) Brightness, as measured by a light meter.
C) Brightness, as measured by people's judgments.
D) Angles as measured in degrees between 0 and 360.
E) Bronze, silver, and gold medals as awarded at the Olympics.
F) Height above sea level.
G) Number of patients in a hospital.
H) ISBN numbers for books.
I) Ability to pass light in terms of the following values: opaque, translucent, transparent.
J) Military rank.
K) Distance from the center of campus.
L) Density of a substance in grams per cubic centimeter.
M) Coat check number.

##### **Answers:**

* Binary, qualitative, ordinal.
* Continuous, quantitative, ratio.
* Discrete, qualitative, ordinal.
* Continuous, quantitative, ratio.
* Discrete, qualitative, ordinal.
* Continuous, quantitative, interval.
* Discrete, quantitative, ratio.
* Discrete, qualitative, nominal.
* Discrete, qualitative, ordinal.
* Discrete, qualitative, ordinal.
* Continuous, quantitative, interval.
* Discrete, quantitative, ratio.
* Discrete, qualitative, nominal.

-----

## Question 3
You are approached by the marketing director of a local company, who believe that he has devised a foolproof way to measure customer satisfaction. He explains his scheme as follows: 

> "It's so simple that I can't believe that no one has thought of it before. I just keep track of the number of customer complaints for each product. I read in a data mining book that counts are ratio attributes, and so, my measure of product satisfaction must be a ratio attribute. But when I rated the products based on my new customer satisfaction measure and showed them to my boss, he told me that I had overlooked the obvious, and that my measure was worthless. I think that he was just mad because our best-selling product had the worst satisfaction since it had the most complaints. Could you help me set him straight?" 

A) **Who is right, the marketing director or his boss? If you answered 'his boss', what would you do to fix the measure of satisfaction?**

* The boss is correct - the marketing director is not taking into account the number of sales of each product. If I have sold 1,000 widgets and 20 whangdoodles, I will almost certainly have a higher number of complaints for the former. It would be better to gauge customer satisfaction using the equation:

$$ Satisfaction_{product} = \frac{n_{complaints}}{n_{sales}} $$

B) **What can you say about the attribute type of the original product satisfaction attribute?**

You can't really say anything about the _attribute type_ of the original measure based on the product satisfaction (see _Example 2_). For example, two products might have the same level of satisfaction from customers, but different numbers of complaints or sales.

-----

## Question 4
A few months later, you are again approached by the same marketing director as in Question 3. This time, he has devised a better approach to measure the extent to which a customer prefers one product over other, similar products. He explains,

> "When we develop new products, we typically create several variations and evaluate which one customers prefer. Our standard procedure is to give our test subjects all of the product variations at one time and then ask them to rank the product variations in order of preference. However, our test subjects are very indecisive, especially when there are more than two products. As a result, testing takes forever. I suggested that we perform the comparisons in pairs and then use the comparisons to get the rankings. Thus, if we have three product variations, we have the customers compare variations 1 and 2, then 2 and 3, and finally 3 and 1. Our testing time with my new procedure is a third of what it was for the old procedure, but the employees conducting the tests complain that they cannot come up with a consistent ranking from the results. And my boss wants the latest product evaluations, yesterday. I should also mention that he was the person who came up with the old product evaluation approach. Can you help me?"

A) **Is the marketing director in trouble? Will his approach work for generating an ordinal ranking of the product variations in terms of customer preference? Explain.**

B) **Is there a way to fix the marketing director's approach? More generally, what can you say about trying to create an ordinal measurement scale based on pairwise comparisons?**

C) **For the original product evaluation scheme, the overall rankings of each product variation are found by computing its average over all test subjects. Comment on whether you think that this is a reasonable approach. What other approaches might you take?**

-----

## Question 5
Can you think of a situation in which identification numbers would be useful for prediction?

Phone numbers could provide a good prediction of owner location (provided that country codes and area codes were included); and student identification numbers could be good predictors of matriculation year, or graduation year.

-----

## Question 6
An educational psychologist wants to use _association analysis_ to analyze test results. The test consists of 100 questions with four possible answers each. 

A) **How would you convert this data into a form suitable for association analysis?**

B) **In particular, what type of attributes would you have an how many of them are there?**

-----

## Question 7
Which of the following quantities is likely to show more temporal autocorrelation: daily rainfall or daily temperature? Why?

-----

## Question 8
Discuss why a document-term matrix is an example of a data set that has _asymmetric discrete_ or _asymmetric continuous_ features.

-----

## Question 9
Many sciences rely on observation instead of (or in addition to) designed experiments. Compares the data quality issues involved in observational science with those of experimental science and data mining.

-----

## Question 10
Discuss the difference between the precision of a measurement and the terms single and double precision, as they are used in computer science, typically to represent floating-point numbers that require 32 and 64 bits, respectively.

-----

## Question 11
Give at least two advantages to working with data stored in text files instead of a binary format.

-----

## Question 12
Distinguish between _noise_ and _outliers_. Be sure to consider the following questions:

A) Is noise ever interesting or desirable? Outliers?
B) Can noise objects be outliers?
C) Are noise objects always outliers?
D) Are outliers always noise objects?
E) Can noise make a typical value into an unusual one, or vice versa?

-----

## Question 13
Consider the problem of finding the $K$ nearest neighbors of a data object. A programmer designs Algorithm 2.1 for this task.

##### Algorithm 2.1: Algorithm for finding K nearest neighbors
1) **for** $i$ = 1 to _number of data objects_ **do**
2)     Find the distances of the $i^{th}$ object to all other objects.
3)     Sort these distances in decreasing order.
4)     **return** the objects associated with the first $K$ distances of the sorted list
5) **end for**

A) **Describe the potential problems with this algorithm if there are duplicate objects in the data set. Assume the distance function will only return a distance of 0 for objects that are the same.**

B) **How would you fix this problem?**

-----

## Question 14
The following attributes are measured for members of a herd of Asian elephants: _weight, height, tusk length, trunk length,_ and _ear area_. Based on these measurements, what sort of similarity measure from Section 2.4 would you use to compare or group these elephants? Justify your answer and explain any special circumstances.

-----

## Question 15
You are given a set of $m$ objects that is divided into $K$ groups, where the $i^{th}$ group is size $m_{i}$. If the goal is to obtaina sample of size $n < m$, what is the difference between the following two sampling schemes? (Assume sampling with replacement.)

A) We randomly select $n*m_{i}/m$ elements from each group.
B) We randomly select $n$ elements from the data set, without regard for the group to which an object belongs.

-----

## Question 16
Consider a document-term matrix, where $tf_{ij}$ is the frequency of the $i^{th}$ word (term) in the $j^{th}$ document and $m$ is the number of documents. Consider the variable transformation that is defined by:

$$ tf'_{ij} = tf_{ij} * \log\frac{m}{df_{i}}, $$
where $df$ is the number of documents in which the $i^{th}$ term appears and is known as the **document frequency** of the term. This transformation is known as the **inverse document frequency** transformation.

A) What is the effect of this transformation if a term occurs in one document? In every document?

B) What might be the purpose of this transformation?

-----

## Question 17
Assume that we apply a square root transformation to a ratio attribute $x$ to obtain the new attribute $x*$. As part of your analysis, you identify an inveral $(a,b)$ in which $x*$ has a linear relationship to another attribute $y$.

A) What is the corresponding interval $(a,b)$ in terms of $x$? 
B) Give an equation that relates $y$ to $x$. In this interval, $y = x^{2}$.

-----

## Question 18
This exercise compares and contrasts some similarity and distance measures.

A) For binary data, the L1 distance corresponds to the Hamming distance; that is, the number of bits that are different between two binary vectors. The Jaccard similarity is a measure of the similarity between two binary vectors. Compute the Hamming distance and the Jaccard similarity between the following two binary vectors.

x = 0101010001
y = 0100011000

B) Which approach, Jaccard or Hamming distance, is more similar to the Simple Matching Coefficient, and which approach is more similar to the cosine measure? Explain.

C) Suppose that you are comparing how similar two organisms of different species are in terms of the number of genes they share. Descrive what measure, Hamming or Jaccard, you think would be more appropriate for comparing the genetic makeup of two organisms. Explain.

D) If you wanted to compare the genetic makeup of two organisms of the same species, e.g., two human beings, would you use the Hamming distance, the Jaccard coefficient, or a different measure of similarity or distance? Explain.

-----

## Question 19
For the following vectors, **x** and **y**, calculate the indicated similarity or distance measures.

A) 

-----

